An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-07-24 poly-embed-journal/poly-embed-journal.tex
Proofreading in the appendix

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\arginf}{arg\,inf}
\DeclareMathOperator*{\sgn}{sgn}

@@

\section{``Book chapter'' approach}
\subsection{Definitions and preliminaries}
\begin{definition}[Closed halfspace]
	For any $w\in\reals^d, b \in \reals$, let $H_{(w,b)}^+ := \{ x \in \reals^d \mid \inprod{x}{w} \geq b\}$ be the closed halfspace defined by $(w,b)$.
\end{definition}
\begin{definition}[Hyperplane]
	For any $(w,b)\in\reals^d \times \reals$, let $H_{(w,b)} := \{ x \in \reals^d \mid \inprod{x}{w} = b\}$ be the hyperplane generated by $(w,b)$, and the boundary of $H^+_{(w,b)}$.
\end{definition}

@@

%\raf{Let's use $\H^*$ for the minimum one, here and below; $\H$ for another one}
%\raf{Similarly $\V^*$ for the minimum one, below; $\V$ for another one}
\begin{theorem}\label{thm:polyhedron-uniquely-gen-facets}
  Given a $d$-dimensional polyhedron $P \subseteq \reals^d$, there is a unique finite set of closed halfspaces $\H^*$ such that $P = \cap \H^*$, and for all finite sets of closed halfspaces $\H$ such that $P = \cap \H$, we have $\H^* \subseteq \H$.
  Moreover, $\{H^+\cap P \mid H^+ \in \H^*\}$ is the set of facets of $P$.
\end{theorem}
\begin{proof}
  \jessie{TODO}
\end{proof}
\jessie{Need to find a good reference.  Cite a result for polytopes and transform the polyhedra into a polytope. Could state the result for class of polyhedra containing $G$.
  Add two facets and now it's a polytope.
}

@@

We now introduce some of our notation.
\begin{itemize}
\item Define $G = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^\Y_+ \times \reals$ to be the hypograph of $g$.
\item We define $H^+_v := H^+_{([v;-1], 0)}$ for a given $v$, where $[v; -1]\in\reals^\Y_+\times\reals$.
\item Additionally, denote $H_y^+ := H_{([e_y; 0], 0)}^+$ for any $y \in \Y$. % \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Similarly, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^\Y_+$ and define $H_y$ analogously. 
\item Define $F_v := H_v \cap G$ for any $v\in\reals^\Y_+$.

@@

\begin{claim}\label{claim:G-intersection-H}
  $G = \cap \H$.
\end{claim}
\begin{proof}
  $(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
  In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$, so $(x,c) \in \cap \H$.
\end{proof}

@@

\begin{claim}\label{claim:G-full-dimensional}
  $G$ is $(|\Y| + 1)$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
  Since $g$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{claim}\label{claim:HY-subset-Hstar}
  $\H_\Y \subseteq \H^*$.
\end{claim}
\begin{proof}
  If there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H^*$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
  The first cannot happen as we take $g$ to be defined (and finite) on $x = e_y \in \reals^\Y_+$.
  Moreover, the second cannot be true by construction of $g$ including the $0-\infty$ indicator on $\reals^\Y_+$.	
\end{proof}

@@

  Since $G$ is full-dimensional, all facets of $G$ are uniquely determined by the hyperplanes $H$ whose halfspaces $H^+$ compose $\H^*$ by Claim~\ref{claim:unique-set-loss-vectors-defining-facets}.
  Any facet $F$ must then be some intersection of an $H_y \cap G$ or $H_v \cap G$.

@@

\begin{claim}\label{claim:g-gen-by-Vstar}
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals_+^\Y)$
\end{claim}
\begin{proof}
  $G = \cap \H = \cap \H^* = \cap (\H_\Y \cup \H_{\V^*}) = \{(x,c) \in \reals_+^\Y \times \reals \mid \inprod{v^*}{x} \geq c$ for all $v^* \in \V^* \}$ where the first equality follows from Corollary~\ref{cor-anything-gen-G-subset-Hstar}.
  Since $G$ is the hypograph of $g$, this means $g$ can be written as 
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals^\Y_+)$. 
\end{proof}

@@

  \jessie{Think of the projection as a linear map and affine hull as a subspace.   Pick any point in relint of the facet, at the linear span, ...}
  \jessie{$\affhull(F_v) = H_v$, $\dim(\affhull(\pi(F_v))) = \dim(\pi(\affhull(F_v)))$.   Show that when you project $H_v$, it gives you $\reals^\Y$.}

  \jessie{Meeting notes: See \url{http://homepages.cae.wisc.edu/~linderot/classes/ie418/lecture11.pdf} for def of AI}
  \jessie{just need to argue that you never needed $c$ (i.e., $c$ was always just $g(x)$).  Maybe: take an affine basis, project, and show it's an affine basis.}
  As a reminder, we define the dimension of a polytope to be the dimension of its affine hull.
  Suppose we are given $|\Y|+1$ affinely independent vectors $z_i$ in $F_v$. 
  We claim their projections $\{\pi(z_i)\}$ are affinely independent.
  Let $a_1 + \ldots + a_{\Y+1} = 0$, such that $\sum_i a_i \pi(z_i) = 0$.
  Want to conclude $a_i = 0$ for all $i$, meaning they are affinely independent.

  Observe $z_i = (x_i, \inprod{v}{x_i})$ for all $i$; therefore, if $z_i \in F_v$ (e.g., $F_v$ supports $G$ at $x, \inprod{v}{x})$), then we also have $z_i \in H_v$.
  So $0 = \sum_i a_i \pi(z_i) = \sum_i a_i x_i$.
  Moreover, the sum $\sum_i a_i z_i = \sum_i a_i (x_i, \inprod{v}{x_i}) = (\sum_i a_i x_i, \inprod{v}{\sum_i a_i x_i}) = (0,0) = \vec 0$.
  Thus, since $a_i = 0$ for all $i$, the set $\{z_i\}$ is affinely independent and the dimensions of the affine hulls are therefore equal.
\end{proof}

\raf{From notes 6/21/2021}
\begin{itemize}
\item REP $\iff$ $\V^* \subseteq \V' := L(\R')$\jessie{Claim~\ref{claim:Vprime-projected-faces-cover-iff-Vstar-subset-Vprime}}
\item REP $\iff \jessie{\Lambda := \{\pi(F_v) \mid v \in \V^*\}; \Lambda \subseteq \{ \pi(F_{v}) \mid v \in L(\R')\}}$.
  \jessie{Claim~\ref{claim:Vprime-projected-faces-cover-iffprojected-faces-subsets}}
\item $\Lambda$ == Full-dim \jessie{Claim~\ref{claim:projected-Vstar-faces-full-dim}.}
\item Lev sets subset $\theta$ \jessie{Just work with $\pi(F_{v})$ where $v \in L(\R)$}
\end{itemize}

@@

\begin{claim}\label{claim:projected-Vstar-faces-full-dim}
  For all $v \in \V^*$, $\pi(F_v)$ is full dimensional in $\reals_+^\Y$.
  \jessie{$\Lambda$ is exactly the set of FDLS; Will lead to $\Theta$ are exactly the full dimensional level sets over simplex (6).}
\end{claim}
\begin{proof}
  By Claim~\ref{claim:unique-set-loss-vectors-defining-facets}, $F_v$ is a facet of $G$ in $\reals^d_+$, meaning it is $((|\Y| + 1) - 1)$-dimensional.
  Moreover, Claim~\ref{claim:pi-preserves-dim} states that the dimension of $F_v$ is preserved for each $v \in \V^*$.
  Thus, $\pi(F_v)$ is $|\Y|$-dimensional.
\end{proof}

@@

\begin{claim}
  $\H_{\V^*} \subseteq \H$
\end{claim}
\begin{proof}
  Observe that it is not immediately clear that $\H_{\V^*} \subseteq \H$, since $\H$ is not finite, and therefore does not immediately form a halfspace representation for $G$.
  If there was a $v^* \in \V^*$ such that $H^+_{v^*} \not \in \H$, then as $\V^*$ generates facets, the projection $\pi(F_{v^*})$ has nonempty interior.
  Fix  $x \in \inter(\pi(F_{v^*}))$; we know $(x,g(x))$ supports $H_v$ at $x$ by Lemma~\ref{lem:projected-faces-iff-support-iff-argmin}.
  Since $L$ is minimizable, we have $\risk L_+ = \min_{v \in \V}\inprod{v}{x} = g(x)$.
  However, since $F_{v^*}$ is a facet \jessie{ref}, it is unique, so $\argmin_{v \in \V}\inprod{v}{x} = \{v^*\}$ and there can be no $v' \in \V$ such that $F_{v'}$ supports $G$ at $x$.
  That is to say that $v^* \not \in \V$, we then contradict the fact that $L$ is minimizable.
  Thus, we must have $\H_{v^*} \subseteq \H$.
\end{proof}

@@

\begin{claim}
  If a $H_v \in \H$ does support $G$, then $F_{v} = H_{v} \cap G$ is a nonempty face of $G$, and thus a subset of a facet. \jessie{Do we use this?}
\end{claim}\label{claim:faces-contained-in-facets}
\begin{proof}
  Since $H_v$ supports $G$, we know that $G \subseteq H_v^+$ and $F_v$ is not empty.
  Since $H^+_v$ is valid, we have $F_v = F_{([v;-1],0)}$ is a face by definition.

@@

\jessie{Don't think this claim is actually necessary?}
\begin{corollary}
	$\Theta = \{\Gamma_r \mid r \in \R^*\}$ 
\end{corollary}

@@

  $\implies$
  This proof follows from three lemmas: first, we observe that $g$ is $1$-homogeneous (Lemma~\ref{lem:g-1-homog}).
  Then we extend the notion of a level set $\Gamma_r$ to the nonnegative orthant $\bar \Gamma_r$, and show that any scalar transformation of a distribution in the level set is contained in the same (extended) level set via Lemma~\ref{lem:levelset-to-extended-levelset}).
  Finally, we show the extended level set is exactly the projection $\pi(F_v)$ (Lemma~\ref{lem:extended-levelset-equals-projected-face}).
  As a corollary, we chain the results to observe $\cup_{r \in \R'} \Gamma_r = \simplex \implies \cup_{r \in \R'} \bar \Gamma_r = \reals_+^\Y = \cup_{v \in L(\R')}\pi(F_v) = \reals^\Y_+$, yielding the forward implication.

>>>>

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\sgn}{sgn}

@@

\section{General characteristics of polyhedral functions}
\subsection{Definitions and preliminaries}
\begin{definition}[Closed halfspace]
	For any $w\in\reals^d, b \in \reals$, let $H_{(w,b)}^+ := \{ x \in \reals^d \mid \inprod{x}{w} \geq b\}$ be the closed halfspace defined by $(w,b)$.
\end{definition}
\begin{definition}[Hyperplane]
	For any $(w,b)\in\reals^d \times \reals$, let $H_{(w,b)} := \{ x \in \reals^d \mid \inprod{x}{w} = b\}$ be the hyperplane generated by $(w,b)$, and the boundary of $H^+_{(w,b)}$.
\end{definition}

@@

%\raf{Let's use $\H^*$ for the minimum one, here and below; $\H$ for another one}
%\raf{Similarly $\V^*$ for the minimum one, below; $\V$ for another one}
\begin{theorem}\label{thm:polyhedron-uniquely-gen-facets}
  Given a $d$-dimensional polyhedron $P \subseteq \reals^d$, there is a unique finite set of closed halfspaces $\H^*$ such that $P = \cap \H^*$, and for all finite sets of closed halfspaces $\H$ such that $P = \cap \H$, we have $\H^* \subseteq \H$.
  Moreover, $\{H^+\cap P \mid H^+ \in \H^*\}$ is the set of facets of $P$.
\end{theorem}
\begin{proof}
  \jessie{TODO} \url{https://arxiv.org/abs/0805.0292} Prop 4.5?
\end{proof}
\jessie{Need to find a good reference.  Cite a result for polytopes and transform the polyhedra into a polytope. Could state the result for class of polyhedra containing $G$.
  Add two facets and now it's a polytope.
}

@@

We now introduce some of our notation.
\begin{itemize}
\item Define $G = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^\Y_+ \times \reals$ to be the hypograph of $g$.
\item We define $H^+_v := H^+_{([v;-1], 0)}$ for a given $v$, where $[v; -1]\in\reals^\Y_+\times\reals$.
\item Additionally, denote $H_y^+ := H_{([e_y; 0], 0)}^+$ for any $y \in \Y$. % \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Similarly, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^\Y_+$ and define $H_y := H_{([e_y;-1], 0)}$. 
\item Define $F_v := H_v \cap G$ for any $v\in\reals^\Y_+$.

@@

\begin{claim}\label{claim:G-intersection-H}
  $G = \cap \H$.
\end{claim}
\begin{proof}
  $(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
  In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$ respectively, so $(x,c) \in \cap \H$.
\end{proof}

@@

\begin{claim}\label{claim:G-full-dimensional}
  $G$ is $d$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
  Since $g$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{claim}\label{claim:HY-subset-Hstar}
  $\H_\Y \subseteq \H^*$.
\end{claim}
\begin{proof}
  If there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H^*$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
  The first cannot happen as we take $g$ to be defined (and finite) on $x = e_y \in \reals^\Y_+$ and is concave.
  Moreover, the second cannot be true by construction of $g$ including the $0-\infty$ indicator on $\reals^\Y_+$.	
\end{proof}

@@

  Since $G$ is full-dimensional, all facets of $G$ are uniquely determined by the hyperplanes $H$ whose halfspaces $H^+$ compose $\H^*$ by Claim~\ref{claim:G-unique-facets-Hstar}.
  Any facet $F$ must then be some intersection of an $H_y \cap G$ or $H_v \cap G$.

@@

We now show that we can equivalently construct $g$ through the unique finite set $\V^*$ instead of a generic set of vectors $\V$.
\begin{claim}\label{claim:g-gen-by-Vstar}
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals_+^\Y)$
\end{claim}
\begin{proof}
  $G = \cap \H = \cap \H^* = \cap (\H_\Y \cup \H_{\V^*}) = \{(x,c) \in \reals_+^\Y \times \reals \mid \inprod{v^*}{x} \geq c$ for all $v^* \in \V^* \}$ where the first equality follows as $\H^* \subseteq \H$.
  Since $G$ is the hypograph of $g$, this means $g$ can be written as 
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals^\Y_+)$. 
\end{proof}

@@

  %\jessie{Think of the projection as a linear map and affine hull as a subspace.   Pick any point in relint of the facet, at the linear span, ...}
  %\jessie{$\affhull(F_v) = H_v$, $\dim(\affhull(\pi(F_v))) = \dim(\pi(\affhull(F_v)))$.   Show that when you project $H_v$, it gives you $\reals^\Y$.}

  \jessie{Meeting notes: See \url{http://homepages.cae.wisc.edu/~linderot/classes/ie418/lecture11.pdf} for def of AI}
  \jessie{just need to argue that you never needed $c$ (i.e., $c$ was always just $g(x)$).  Maybe: take an affine basis, project, and show it's an affine basis.}
  As a reminder, we define the dimension of a polytope to be the dimension of its affine hull.
  Suppose we are given $|\Y|+1$ affinely independent vectors $z_i$ in $F_v$. 
  We claim their projections $\{\pi(z_i)\}$ are affinely independent.
  Let $a_1 + \ldots + a_{\Y+1} = 0$, such that $\sum_i a_i \pi(z_i) = 0$.
  We want to conclude that we must have $a_i = 0$ for all $i$, meaning they are affinely independent.

  Observe $z_i = (x_i, \inprod{v}{x_i})$ for all $i$; therefore, if $z_i \in F_v$ (e.g., $F_v$ supports $G$ at $x, \inprod{v}{x})$), then we also have $z_i \in H_v$.
  So $0 = \sum_i a_i \pi(z_i) = \sum_i a_i x_i$.
  Moreover, the sum $\sum_i a_i z_i = \sum_i a_i (x_i, \inprod{v}{x_i}) = (\sum_i a_i x_i, \inprod{v}{\sum_i a_i x_i}) = (\vec 0,0) = \vec 0$.
  Thus, since $a_i = 0$ for all $i$, the set $\{z_i\}$ is affinely independent and the dimensions of the affine hulls are therefore equal.
\end{proof}

\raf{From notes 6/21/2021}
\jessie{Where we're going: (can be deleted)}
\begin{itemize}
\item REP $\iff$ $\V^* \subseteq \V' := L(\R')$\jessie{Claim~\ref{claim:Vprime-projected-faces-cover-iff-Vstar-subset-Vprime}}
\item REP $\iff \jessie{\Lambda := \{\pi(F_v) \mid v \in \V^*\}; \Lambda \subseteq \{ \pi(F_{v}) \mid v \in L(\R')\}}$.
  \jessie{Claim~\ref{claim:Vprime-projected-faces-cover-iffprojected-faces-subsets}}
\item $\Lambda$ == Full-dim \jessie{Claim~\ref{claim:projected-Vstar-faces-full-dim}.}
\item Lev sets subset $\theta$ \jessie{Just work with $\pi(F_{v})$ where $v \in L(\R)$}
\end{itemize}

@@

\begin{claim}\label{claim:projected-Vstar-faces-full-dim}
  For all $v \in \V^*$, $\pi(F_v)$ is full dimensional in $\reals_+^\Y$.
  \jessie{$\Lambda$ is exactly the set of FDLS; Will lead to $\Theta$ are exactly the full dimensional level sets over simplex (6).}
\end{claim}
\begin{proof}
  By Claim~\ref{claim:unique-set-loss-vectors-defining-facets}, $F_v$ is a facet of $G$ in $\reals^d_+$, meaning it is $(d - 1)$-dimensional.
  Moreover, Claim~\ref{claim:pi-preserves-dim} states that the dimension of $F_v$ is preserved for each $v \in \V^*$.
  Thus, $\pi(F_v)$ is $|\Y|$-dimensional.
\end{proof}

@@

\begin{claim}
  $\H_{\V^*} \subseteq \H$
\end{claim}
\begin{proof}
  Observe that it is not immediately clear that $\H_{\V^*} \subseteq \H$, since $\H$ is not finite, and therefore does not immediately form a halfspace representation for $G$.
  If there was a $v^* \in \V^*$ such that $H^+_{v^*} \not \in \H$, then as $\V^*$ generates facets, the projection $\pi(F_{v^*})$ has nonempty interior.
  Fix  $x \in \inter(\pi(F_{v^*}))$; we know $(x,g(x))$ supports $H_v$ at $x$ by Lemma~\ref{lem:projected-faces-iff-support-iff-argmin}.
  Since $L$ is minimizable, we have $\risk L_+ = \min_{v \in \V}\inprod{v}{x} = g(x)$.
  However, since $F_{v^*}$ is a facet \jessie{ref}, it is unique, so $\argmin_{v \in \V^*}\inprod{v}{x} = \{v^*\}$ and there can be no $v' \in \V^*$ such that $F_{v'}$ supports $G$ at $x$.
  That is to say that $v^* \not \in \V$, we then contradict the fact that $L$ is minimizable.
  Thus, we must have $\H_{v^*} \subseteq \H$.
\end{proof}

@@

\begin{claim}\label{claim:faces-contained-in-facets}
  If a $H_v \in \H$ does support $G$, then $F_{v} = H_{v} \cap G$ is a nonempty face of $G$, and thus a subset of a facet. \jessie{Do we use this?}
\end{claim}
\begin{proof}
  Since $H_v$ supports $G$, we know that $G \subseteq H_v^+$ and $F_v$ is not empty.
  Since $H^+_v$ is valid, we have $F_v = F_{([v;-1],0)}$ is a face by definition.

@@

\begin{corollary}
	$\Theta = \{\Gamma_r \mid r \in \R^*\}$ 
\end{corollary}

@@

  $\implies$
  This proof follows from three lemmas: first, we observe that $g$ is $1$-homogeneous (Lemma~\ref{lem:g-1-homog}).
  Then we extend the notion of a level set $\Gamma_r$ to the nonnegative orthant $\bar \Gamma_r$, and show that any scalar transformation of a distribution in the level set is contained in the same (extended) level set via Lemma~\ref{lem:levelset-to-extended-levelset}.
  Finally, we show the extended level set is exactly the projection $\pi(F_v)$ (Lemma~\ref{lem:extended-levelset-equals-projected-face}).
  As a corollary, we chain the results to observe $\cup_{r \in \R'} \Gamma_r = \simplex \implies \cup_{r \in \R'} \bar \Gamma_r = \reals_+^\Y = \cup_{v \in L(\R')}\pi(F_v) = \reals^\Y_+$, yielding the forward implication.
